The challenges of teaching cybersecurity as part of mainstream computer science and related disciplines?

### 2019-10-08 CEP20/CEP_cybersecurity.tex
Early draft

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{How well do we teach Cybersecurity}

@@

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Crick, Davenport,  Irons, and Prickett.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. Based on the ``acmart'' document class, this
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002978</concept_id>
<concept_desc>Security and privacy</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003527.10003529</concept_id>
<concept_desc>Social and professional topics~Accreditation</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

@@

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Accreditation}

@@

\section {What is it?}
A short description of the practice you're presenting

@@

@@

\section {Does it work?}	How do you know? Give some evidence of effectiveness in context.
\section {Who else has done this?}	Where did you get the idea from? (If from published reports, please include references). How did you find out about it? Was it easy/hard to adopt? What did you change?
\section {What will you do next?}	Will you vary this, or develop it further?
\section{Why are you telling us this?}	What is interesting, or useful, about this to someone else?

>>>>

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The challenges of teaching cybersecurity as part of mainstream computer science and related disciplines?}

@@

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Crick, Davenport,  Irons, and Prickett.}
%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  "Every Computer Science Degree Should Require a Course in Cybersecurity" **TO DO add reference**
  argued a recent article in the Harvard Business review. Computer Science and related degrees in the united Kingdom are evolving to meet this expectation. The BCS, The Charterted Institute for IT has been mandating this for a number of years \cite{Cricketal2019}. This may have encouraged these developments, although the discipline has been embedding cybersecurity before it was mandated by the BCS. Delivering cybersecurity effectively presents a number of challenges related to research and pedagogy. This article explores the progress to date, challenges for the future and highlights areas of potential  enhancement.

\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002978</concept_id>
<concept_desc>Security and privacy</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003527.10003529</concept_id>
<concept_desc>Social and professional topics~Accreditation</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

@@

% To do add cybersecurity to this

%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Accreditation}

@@

\section {What is it?}
%A short description of the practice you're presenting

This paper explores the challenges related to research and pedagogy of delivering the this cybersecurity content.

@@

Cybersecurity is becoming increasing pivotal to the operation of organisations of all sizes and all organisations are increasingly expected to make reasonable adjustments to protect their activities

\begin{quote}
	``{\emph{...[need to] change the culture in your organisation around cyber security; to try to do for cyber what has been done so successfully for health and safety, for example, over the last ten years --- to get everybody to take it seriously; to take the risk management process seriously and drive that down through the organisation.}}''\\
	\hfill Robert Hannigan~\cite{Hannigan2019a}, former Director of GCHQ
\end{quote}

This focus on cybersecurity, includes calls for formal education -- school-level as well as tertiary -- to respond to this situation, at the individual level and via recommended curricula~\cite{mcgettrick-et-al:sigcse2014,ACM2017b} and professional accreditation requirements~\cite{BCS2018a,NCSC2017}. This is further reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK~\cite{brown-et-al:toce2014,murphy-et-al:programming2017,tryfonas+crick:petra2018}

Computer Science and related disciplines are evolving to meet these demands. However doing so is not without challenges. This paper explores the progress to date, challenges for the future and highlights a number of enhancement activities for the discipline.

@@

\section {Does it work?}	How do you know? Give some evidence of effectiveness in context.

The UK situation appears relatively advanced compared to other jurisdictions. 61\% of UK courses offer mandatory cybersecurity content, and this research was based on web scraping~\cite[Table 1]{Ruiz2019a}. As such it represents a lower bound since not all coverage will necessarily be clearly articulated in publicly available documentation online.

In many jurisdictions, Professional, Statutory and Regulatory Bodies (PSRBs) have responded to these expectations by adjusting their accreditation requirements. In the UK, BCS, The Chartered Institute for IT (BCS) has had a requirement to include information security in the curriculum since 2010, and has expected coverage of an agreed cybersecurity syllabus since 2015, with the result that all accredited universities should be compliant by 2020 (due to the five-year cycle). More precisely, accredited degrees have been expected to demonstrate coverage of ``{\emph{2.1.9 Knowledge and understanding of information security issues in relation to the design, development and the use of information systems}}'' \cite[p.~30]{BCS2018a} since 2010 with an enhanced cybersecurity related definition of what this entails since 2015 \cite[p.~17--18]{BCS2018a}.

It has been reported that this initiative is showing good progress \cite{Cricketal2019} with the majority university's already meeting these expectations and a further number evolving their provision so that they do so.

Collective challenges still persist related to underpinning resources, staffing and expected pedagogies.

\section {Who else has done this?}	Where did you get the idea from? (If from published reports, please include references). How did you find out about it? Was it easy/hard to adopt? What did you change?

In jurisdictions other than the UK, PSRB's have also adjusted their expectations to enhance the cybersecurity provision. For example in the United States of America (USA), the Association of Computing Machinery (ACM) has equally had cybersecurity (IAS: ``Information Assurance and Security'') in the curriculum since 2013~\cite{ACM2013a}, but it is not the accrediting body. The Accreditation Board for Engineering and Technology (ABET) is, and is requiring IAS with effect from the 2019-20 cycle (self-study reports due 1 July 2019): more precisely \cite[Table 3]{Oudshoornetal2018a} ``{\emph{The computing topics must include: \dots{} Principles and practices for secure computing\dots}}''. This should mean that  all accredited universities should be compliant by 2025 (due to their six-year cycle)

The ACM/IEEE-CS Joint Task Force on Computing Curricula~\cite[p.~97]{ACM2013a} takes a distinct view on the Knowledge Areas (KAs):

\begin{quote}
	``{\emph{The Information Assurance and Security KA is unique among the set of KAs presented here
			given the manner in which the topics are pervasive throughout other KAs.}}''
\end{quote}

It proposes nine ``core'' hours and 63.5 distributed across the other KAs. Nevertheless, the situation on the ground in the USA is different~\cite{Ackerman2019a}:

\begin{quote}
	``{\emph{Universities suffer shortcomings, as well. Roughly 85 of them offer undergraduate and/or graduate degrees in cybersecurity. There is a big catch, however. Far more diversified computer science programs, which attract substantially more students, don't mandate even one cybersecurity course.}}''
\end{quote}

However, whilst the accreditation regimes in the UK are mandating inclusion of cybersecurity there remain challenges in terms of the underpinning resources, the available Faculty resources and pedagogies required to effectively deliver the cybersecurity content.

\section {What will you do next?}	Will you vary this, or develop it further?
\section{Why are you telling us this?}	What is interesting, or useful, about this to someone else?
